---
title: "Module2_Assignment_YO"
author: "Yuki Obayashi"
date: "2025-01-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading and preprocessing the data

```{r, echo = TRUE}
library(readr)
activity <- read_csv("activity.csv")
```

# What is mean total number of steps taken per day?

## Calculate the total number of steps taken per day
```{r, echo = TRUE}
library(tidyverse)
total_steps_per_date <- activity %>% 
  group_by(date) %>% 
  summarize(total_steps = sum(steps, na.rm = TRUE))

print(total_steps_per_date)

```
## Make a histogram of the total number of steps taken each day
The main difference between histogram and bar graph is whether the x axis is continuous or categorical.
In this assignment, the x axis is date. Although date is a continuous variable, it should be categorized as categorical in R plotting. Therefore, I used geom_bar() in the following code.

```{r, echo = TRUE}
library(dplyr)
library(ggplot2)

ggplot(total_steps_per_date, aes(x = date, y = total_steps)) +
  geom_bar(stat = "identity", fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Total Steps Per Day", x = "Date", y = "Total Steps") +
  theme_minimal()

```

## Calculate and report the mean and median of the total number of steps taken per day

```{r, echo = TRUE}
activity %>% 
  group_by(date) %>% 
  summarize(total_steps_mean = mean(steps, na.rm = TRUE),
            total_steps_median = median(steps, na.rm = TRUE))

```

# What is the average daily activity pattern?

## Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
activity_time_series <- activity %>%
  group_by(interval) %>%
  summarize(average_steps = mean(steps, na.rm = TRUE))

ggplot(activity_time_series, aes(x = interval, y = average_steps)) +
  geom_line() +  
  labs(title = "Average Daily Activity Pattern",
       x = "5-Minute Interval", 
       y = "Average Number of Steps") +
  theme_minimal()

```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
activity_time_series %>%
  filter(average_steps == max(average_steps, na.rm = TRUE))

```
The answer is 8:35.


# Imputing missing values

## Calculate and report the total number of missing values in the dataset

```{r}
missing_rows <- activity %>% filter(is.na(steps))
nrow(missing_rows)
```

## Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated.
In addition to using the mean/median for that day, or the mean for that 5-minute interval, other option includes to use preceding or succeeding values.

## Create a new dataset that is equal to the original dataset but with the missing data filled in.
Here, I try to fill in the missing values using the mean for that day.
If all the steps in that day are missing, I filled it as 0.

```{r}

activity_filled <- activity %>% 
  group_by(date) %>% 
  mutate(steps = ifelse(is.na(steps), coalesce(mean(steps, na.rm = TRUE), 0), steps))
print(activity_filled)

```

## Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}

total_steps_per_date_filled <- activity_filled %>% 
  group_by(date) %>% 
  summarize(total_steps = sum(steps, na.rm = TRUE))

#Histgram
ggplot(total_steps_per_date_filled, aes(x = date, y = total_steps)) +
  geom_bar(stat = "identity", fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Total Steps Per Day", x = "Date", y = "Total Steps") +
  theme_minimal()

#Mean and median
activity_filled %>% 
  group_by(date) %>% 
  summarize(total_steps_mean = mean(steps, na.rm = TRUE),
            total_steps_median = median(steps, na.rm = TRUE))


```

As long as I adopt the strategy to fill the missing values in 0, the results are similar.

# Are there differences in activity patterns between weekdays and weekends?

## Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}

activity_weekday <- activity %>% 
  mutate(weekday = weekdays(date)) %>% 
  mutate(weekday_or_end = ifelse(weekday %in% c("Saturday", "Sunday"), "weekend", "weekday"))

print(activity_weekday)


```

## Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
#weekday
activity_time_series_weekday <- activity_weekday %>%
  filter(weekday_or_end=="weekday") %>% 
  group_by(interval) %>%
  summarize(average_steps = mean(steps, na.rm = TRUE))

plot_weekday <- ggplot(activity_time_series_weekday, aes(x = interval, y = average_steps)) +
  geom_line() +  
  labs(title = "Average Daily Activity Pattern (weekday)",
       x = "5-Minute Interval", 
       y = "Average Number of Steps") +
  theme_minimal()

#weekend
activity_time_series_weekend <- activity_weekday %>%
  filter(weekday_or_end=="weekend") %>% 
  group_by(interval) %>%
  summarize(average_steps = mean(steps, na.rm = TRUE))

plot_weekend <- ggplot(activity_time_series_weekend, aes(x = interval, y = average_steps)) +
  geom_line() +  
  labs(title = "Average Daily Activity Pattern (weekend)",
       x = "5-Minute Interval", 
       y = "Average Number of Steps") +
  theme_minimal()

#Arrange
library(gridExtra)
grid.arrange(plot_weekday, plot_weekend, ncol = 2)

```

In weekends, the total number of steps increases in the afternoon.
